---
title: "Proportion Sheet Dataset"
author: "Eden Barker"
date: "25/02/2022"
output:
  pdf_document
toc: FALSE
bibliography: references.bib
---

```{r}
knitr::opts_chunk$set(echo = FALSE, include = FALSE)
install.packages("readr")
install.packages("knitr")
install.packages("kableExtra")
library(readr)
library(janitor)
library(tidyverse)
library(tidyr)
library(dplyr)
library(knitr)
library(kableExtra)
```

```{r}
#read dataset
GSS_Table3_Proportion <- read.csv("~/Australia_Stats/inputs/GSS_Table3_Proportion.csv")
View(GSS_Table3_Proportion)

#clean dataset
groups_df<- 
  clean_names(GSS_Table3_Proportion)
groups_df

groups_df <-
  select(x2,x3,x5,x6,x8,x9) |>
  slice(29,152)
  view(groups_df)
```

```{r data exploring, echo =FALSE}
### Table ###
groups_df <- 
  tibble(Gender = c('Male', 'Female'),
         with_MHI = c('21', '20'),
         without_MHI = c('13','10'),
         with_LHI = c('19','17'),
         without_LHI = c('11','7'),
         with_LD = c('20','16'),
         without_LD = c('13','8'))
view(groups_df)

#Figure3
kbl(groups_df) %>% 
  kable_classic() %>%
  add_header_above(c(" " = 1, "Mental Health Condition" = 2, "Long-Term Health Condition" = 2, "Long-Term Disability" = 2))

```
#Data

## Data Cleaning
  The data cleaning was processed on an opensource platform called 'R'[@citeR]. The dataset was pulled from the **Australian Bureau of Statistics** as a downloadable xlxs file with over 377 observations and 10 variables that has already been cleaned and tallied, separated by gender outcomes. The focus of this dataset was to provide insight on relationships between characteristics. Due to the nature of the dataset being a results summary that is meant for public consumption, we understand that it can be viewed and analyzed to serve many purposes. For the purpose of having a meaningful analysis,we controlled the dataset down to main variables which are the health conditions and gender of participants. We decided to have a focal point that we can concentrate on, rather than interpreting the dataset as whole. This allows us to substantiate the findings and postulate recommendations based from our own analysis. The data has been cleaned using the package 'Janitor' [@Janitor] and filtered to create a Tibble using packages 'Dplyr' [@Dplyr], 'TidyR' [@TidyR], 'Tidyverse' [@Tidyverse] and 'Readr' [@ReadR]. The tibble was generated to focus on data for the *Cared for a person with either a disability, long term health condition or old age in last 4 weeks* entry with a focus on outcomes from **Males** and **Females**.[@ABS_GSS2020] The dataset was reduced to only include reports for **Mental Health Disabilities, Long-Term Health Disabilities and Long-Term Disabilities**. 

##Data Source
  The dataset was generated by the **Australian Bureau of Statistics** called *General Social Survey: Summary Results Australia 2020* [@ABS_GSS2020] It is an annual survey that aims to provide information on "social characteristics, wellbeing and social experiences of people in Australia". The results summary was released on June 29, 2021, referencing the period 2020 as the year the data was collected.[@ABS_GSS2020] The **Australian Bureau of Statistics** indicated that the survey was conducted over a 4 month period from 15th June to 5th September 2020 during the COVID-19 pandemic. As per **Australian Bureau of Statistics** "There was no face-to-face interviewing conducted in 2020 due to COVID-19 restrictions. Because of this changed methodology and the impact of COVID-19 restrictions on the Australian population, care should be exercised when making comparisons with 2019.The survey was previously ran in 2019.  Prior to that, the survey was conducted once every four years from 2002 to 2014". [@ABS_GSS2020] However we did not compare data from 2019 on this report and focused on an aspect we thought to be an occurrence that may have come about during the pandemic. The dataset was pulled from the **Australian Bureau of Statistics** database where it was downloaded locally as an xlxs file and converted to a csv file. It was then processed in 'R' [@citeR] and visualized as graphs and table with a controlled dataset using packages 'Ggplot2' [@Ggplot2] and 'KableExtra' [@KableExtra]. 

###Methodology 
  The *General Social Survey: Summary Results Australia 2020* data includes all Australian residents aged 15 and over. There were approximately 5,300 Australian households where data was collected from and were completed online or via telephone interview. [@ABS_GSS2020] The key benefit of this survey is to provide information on the multi-dimensional nature of relative advantage and disadvantage across the population. 

# Discussion

## Secondary Research
McGillvey and Dodevska posits Australia to be a country that values volunteering as one of its core cultural feature @McVillyandDodevska2017. This translates to the dataset presented by the Australian Bureau of Statistics’ summary results of the General Social Survey of 2020. @ABS_GSS2020 Focusing on the specific results regarding volunteering, it’s remarkable that the proportion of volunteers between those who suffer from conditions of *Mental Health Disabilities, Long-Term Health Disabilities and Long-Term Disabilities* to those who don’t are more than half. @ABS_GSS2020 According to Australian Institute of Health and Wellness @AIHW_2021, volunteering serves as an indicator of wellbeing and social cohesion of people. Through volunteering, people can become more outwardly focused, which may strengthen social interactions or even reduce anxiety. @AIHW_2021 As shown in the table **Figure 3**, we see that there are less male and female volunteers who are not suffering from any conditions such as *Mental Health Disabilities, Long-Term Health Disabilities and Long-Term Disabilities*. Interestingly the data shows that there are slightly more male volunteers than females. This can suggest that this information may be informally setup such as unpaid work or support for a family member by a family member.gi

To be continued by Eden: 
arguments : 
 1 talk about the possibility of less funding hence family members may be needing to support disabled family as informal carers due to lack of financial support McGillvey 
 2 people with disabilities may meet the people with same disabilities leading to marriage and inadvertently becoming each other support/carer
 3 these surveys are meant to create better policies for the community, due to decrease in volunteers it may have a positive effect in term of increasing funding for volunteer programs... ? 
 4 cite the article about covid possibily becoming a long term disability and how an ongoing pandemic for 2 years may exacerbate the need for more govt support, financially etc and volunteers

\newpage

\appendix

\newpage


# References

# Appendix {-}

